---
title: "Executing a Basic GOR Query"
author: "Edvald Gislason <edvald@wuxinextcode.com>"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Basic GOR Query}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
requireNamespace("knitr")
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

In this example, we're going to run a simple remote gor query using the `gorr` package. 

## Load packages

First load the `gorr` package, the `tidyverse` package is recommended in general, but not required for this example

```{r packages}
library(gorr)
library(magrittr) # pipe
library(tibble) 
```

## Connecting to the direct query service

First we'll need to establish a connection to our direct query API. To do that we'll need to call `gor_connect` and provide it with the relevant parameters pointing to the direct-query-service, i.e. `api_key` and `project`:

```{r}
api_key <- Sys.getenv("GOR_API_KEY")
conn <- gor_connect(api_key, project = Sys.getenv("GOR_API_PROJECT"))
conn
```

If everything goes as planned, we'll have a `conn` object to 
pass into the `gor_query` function to finally run a query:

```{r}
result <- gor_query("gor #dbsnp# | top 100", conn)
```

The results come back as an R `data.frame`:


```{r}
print(result)
```

It should also be noted that larger gor queries can be constructed inside a string block and piped directly into the 
`gor_query` function:

```{r}
chr21_results <- "
  gor -p chr21 #dbsnp# 
    | where len(reference)=1 and len(allele)=1
    | calc snptype reference+'/'+allele 
    | hide rsIDs
    | top 100" %>%
    gor_query(conn)
```

```{r}
print(chr21_results)
```
