---
title: "Covariate services"
author: "Andri M. Stefansson <andri@genuitysci.com>"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Covariate services}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
requireNamespace("knitr")
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


In this example, we're going to demonstrate how to interact with the covariate services API using the `gorr` package.

## Loading required packages

```{r packages}
library(gorr)
library(magrittr) # pipe
```

## Connect to Genuity Science's services.

First we'll need to establish a connection to our API services. To do that we'll need to call `platform_connect` and provide it with the relevant parameters pointing to the phenotype-catalog-service, i.e. `api_key` and `project`:


```{r}
conn <- platform_connect(api_key = Sys.getenv("GOR_API_KEY"),
                          project = Sys.getenv("GOR_API_PROJECT"))
conn
```

If everything goes as planned, we'll have a `conn` object to pass into subsequent functions.

## List available covariates

Covariates available to users can be listed by passing the `conn` object to `get_covariates`.

```{r, eval=FALSE}
covariates <- get_covariates(conn)
covariates
```

The results come back as a `list` of covariates available to the user.


## Get phenotype

Now can fetch a covariate from the project `get_covariate` and passing the covariate's unique `id` along with the `conn` object .

```{r, eval=FALSE}
covariate <- get_covariate(id = 1, conn)
covariate
```
