---
title: "Using create statements and referencing local data frames"
author: "Edvald Gislason <edvald@wuxinextcode.com>"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{GOR Create Statements and Virtual Relations}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
requireNamespace("knitr")
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

In this example, we're going to show how the `gor_create` can be used to prepare and construct a query <a href="https://en.wikipedia.org/wiki/Closure_(computer_programming)">closure</a>. This both reduces repetitions in code, as well as simplifies iterative workflows in GOR. 

## Load packages

First load the `gorr` package, the `tidyverse` package is recommended:

```{r packages}
library(gorr)
suppressPackageStartupMessages(library(tidyverse))
```

Next we make a `conn` object for holding information on the API we're connecting to. `gor_connect` takes 2 parameters `api_key` and `project` but if either are left out then it will try to read the environment variables `GOR_API_KEY`, and `GOR_API_PROJECT` respectively. Here below we have the `GOR_API_KEY` environment variable already defined so supplying the function only with a target project suffices. After this we create a `query` function/closure so we don't have to reference `conn` again:

```{r}
conn <- gor_connect(project = "ukbb_hg38")
query <- gor_create(conn = conn)
query
```

Now we can call that function with only the query parameter. Let's search for genes containing _BRCA_ and save the resulting table as a local dataframe `mygenes`:

```{r}
mygenes <- query("gor #genes# | grep BRCA")
mygenes
```

Next we can expand on our previously defined `query` function by supplying it back into `gor_create` as the `replace` parameter. This time we include some definitions using the `defs` parameter and then we can alias our local dataframe so that we can reference it in remote queries. In GOR this is called _virtual relations_:

```{r}
query <- gor_create(
  defs = "def variants = #dbsnp#",
  mygenes = mygenes, 
  replace = query
)
query
```
Now that we have our updated `query` function, we can use it to `gor` our table of genes and join it to the `#dbsnp#` table we aliased as `variants` in the definitions part above. The result is a list of all variants within each gene in our table

```{r}
brca_variants <- query("
    gor [mygenes] | join -segvar variants        
")

brca_variants %>% 
  group_by(gene_symbol) %>%
  summarize(records = n(),
            variants = n_distinct(rsids))
```
The reason for the difference in `# records `and `# variants` above can be explained by looking into the data:

```{r}
target_variant <- 
  brca_variants %>%
  group_by(rsids) %>%
  count() %>% 
  ungroup() %>%
  arrange(desc(n)) %>%
  head(n = 1) %>%
  pull(rsids)

target_variant
```


```{r}
brca_variants %>% filter(rsids == target_variant) %>% select(-distance)
```

