---
title: "Phenotype services"
author: "Andri M. Stefansson <andri@genuitysci.com>"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Phenotype services}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
requireNamespace("knitr")
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

In this example, we're going to explore the capabilities of the phenotype-services using the `gorr` package. 

## Load packages

First load the `gorr` package, the `tidyverse` package is recommended in general, but not required for this example

```{r packages}
library(gorr)
library(magrittr) # pipe
library(tibble) 
```

## Connect to Genuity Science's services.

First we'll need to establish a connection to our API services. To do that we'll need to call `platform_connect` and provide it with the relevant parameters pointing to the phenotype-catalog-service, i.e. `api_key` and `project`:


```{r}
conn <- platform_connect(api_key = Sys.getenv("GOR_API_KEY"),
                          project = Sys.getenv("GOR_API_PROJECT"))
conn
```

If everything goes as planned, we'll have a `conn` object to pass into subsequent functions.


## List project phenotypes

Let's start by fethcing available phenotypes in project (first 25)

```{r}
phenotypes <- get_phenotypes(conn, limit=25)
print(phenotypes[1:4])
```

The results come back as a `list` of phenotypes

## Get phenotype

Next, let's fetch a phenotype from the project. We'll use the first listed 

```{r}
phenotype <- get_phenotype(name = names(phenotypes)[1], conn)
phenotype
```

The results come back as a `phenotype` object, which is a list of lists containing different info on the `phenotype` object.

## Create phenotype

We can add a new phenotype to the project using the `create_phenotype` function as follows

```{r}
name <- paste0("rtest_pheno", sample(1:99,1)) # Name of new phenotype
result_type <- "CATEGORY" # Type of phenotype (either "QT", "SET" or "CATEGORY")
description <- "This is a test phenotype" # Optional phenotype description
new_phenotype <- create_phenotype(name, result_type, conn, description)
new_phenotype
```

A new phenotype has now been added to the project's phenotype-catalog as well as been assigned to the variable `new_phenotype`.

Phenotypes can also be created using the full set of metadata in a single call and with a GOR/NOR query.

```{r, eval=FALSE}
##NOT RUN##
gor_query <- "
def #phenos# = added_salt_to_food;
create ##aggregate## = nor UKBB/phenotypes/fields.nord -s phenotype -f #phenos#
    | ATMAX visit_id -gc PN,phenotype
    | map -h -m '#empty#' -c phenotype,value <(nor -h UKBB/phenotypes/meta/field_encoding_lookup.tsv)
    | REPLACE value if(meaning = '#empty#',value,meaning)
    | hide meaning,visit_id,index_id;
nor [##aggregate##] | sort -c PN
    | pivot phenotype -ordered -gc PN -v #phenos# -e NA
    | rename (.*)_value #{1}
    | select PN
"

name2 <-  paste0("rtest_pheno", sample(1:99,1))
description2 <- "A set of individuals who answered YES to the question: do you salt your food? https://biobank.ndph.ox.ac.uk/ukb/field.cgi?id=104660"
result_type2 <- "SET"
category2 <- "symptoms"
query2 <- gor_query
tags2 <- "dietary,nutrition"

new_phenotype2 <- create_phenotype(name2, result_type2, conn, description2, query = query2, tags = tags2)
new_phenotype2
phenotype_delete(new_phenotype2, conn)
```

If we see the phenotype's state is 'pending', or we simply want to refresh the phenotype for fetching most recent info on the phenotype,  
we can refresh the phenotype using the `phenotype_refresh` method

```{r}
new_phenotype <- phenotype_refresh(new_phenotype, conn)
new_phenotype
```

In case the query fails we can call either `phenotype_get_error(phenotype, conn)`, returning latest error (formatted) or `phenotype_get_errors(phenotype, conn)` which returns all errors as lists along with their timestamps. 

## Update phenotype description

Passing a phenotype object and a new description to `phenotype_update_description`  updates the phenotype's description in the project's phenotype-catalog. The output is, again, an updated `phenotype` object.

```{r}
new_description <- "An updated description"
new_phenotype <- phenotype_update_description(new_description, new_phenotype, conn)
new_phenotype
```

## Upload phenotype data

Let's start by creating a dummy dataset which is a list of list:
[[PN#1, attribute], [PN#2, attribute], ... , [PN#N, attribute]]

```{r}
cohort_size <- 2000
pns <- as.character(1000:cohort_size+1000)
data <- lapply(pns, function(x) { list(x, if (runif(1)>0.5) 'obease' else 'lean') })
print(data[1:2])

phenotype_upload_data(new_phenotype, data, conn)

```
## Get phenotype data

Phenotype data can then be fetched using `get_data`:

```{r}
get_data(new_phenotype, conn)
```

## Get all tags available in the phenotype catalog

We can list all available tags in the catalog using `get_tags` and passing the `platform_connection` object.

```{r}
get_tags(conn)[1:4]
```

## Get phenotype tags

Retrieve all tags for phenotype

```{r}
phenotype_get_tags(new_phenotype)
```

## Add phenotype tag/s

Add a new tag/s to this phenotype. Multiple tags should be separated by either comma separated string, `"tag1,tag2"`, character vector `c("tag1", "tag2")` or a combination of the two

```{r}
new_phenotype <- phenotype_add_tag(tag = "testTag2", new_phenotype, conn)
phenotype_get_tags(new_phenotype)
```

for assurance we can get the phenotype from server and check tags

```{r}
get_phenotype(name, conn) %>% phenotype_get_tags()
```

## Set phenotype tag/s

Set the tag list for this phenotype, overriding all previous tags, defining the tags adheres to the same rules as add_phenotype_tag

```{r}
tags <- "testTag7,testTag8"
new_phenotype <- phenotype_set_tags(tags, new_phenotype, conn)
phenotype_get_tags(new_phenotype)

```

again, for assurance let's load the tags from server

```{r}
get_phenotype(name, conn) %>% phenotype_get_tags()
```

## Delete phenotype tag/s

Delete a tag/s from phenotype.

```{r}
new_phenotype <- phenotype_delete_tag(tag = "testTag7", new_phenotype, conn)
phenotype_get_tags(new_phenotype)
#or get_phenotype(name, conn) %>% phenotype_get_tags()
```


## Delete phenotype

A phenotype can easily be removed from a project using the `phenotype_delete` function.
`phenotype_delete` expects a `phenotype` object. Therefore the phenotype needs to be fetched using the `get_phenotype` function if it has not been assigned to a variable already.

```{r}
phenotype_delete(new_phenotype, conn)
```
