---
title: "Phenotype playlists"
author: "Andri M. Stefansson <andri@genuitysci.com>"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Phenotype playlists}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
requireNamespace("knitr")
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


In this example, we're going to demonstrate how to interact with the phenotype playlist API using the `gorr` package.

One or more phenotypes can be grouped into a playlist. A playlist works like an added layer of curation on the phenotypes. For example, it could be a set of heart failure-related phenotypes or a set of principle components. It may just be an easy way to share work between people working on a team. Playlists make it easier to recall work between days. The methods for working with playlists are below:

## Loading required packages

```{r packages}
library(gorr)
library(magrittr) # pipe
```

## Connect to Genuity Science's services.

First we'll need to establish a connection to our API services. To do that we'll need to call `platform_connect` and provide it with the relevant parameters pointing to the phenotype-catalog-service, i.e. `api_key` and `project`:


```{r}
conn <- platform_connect(api_key = Sys.getenv("GOR_API_KEY"),
                          project = Sys.getenv("GOR_API_PROJECT"))
conn
```

If everything goes as planned, we'll have a `conn` object to pass into subsequent functions.


## List available playlists

Playlists available to users can be listed by passing the `conn` object to `get_playlists`.

```{r}
playlists <- get_playlists(conn)
playlists
```

The results come back as a `vector` of playlist names available to the user.


## Create playlist

If we want to add a new playlist to the catalog we can do so using `create_playlist` along with the name of the playlist and optionally description and name of phenotypes to be addied to the playlist.
To add multiple phenotypes when initializing a new playlist you can either do so by providing a comma seperated string, `"pheno1,pheno2"` or character vector `c("pheno1","pheno2")`

NB: Before adding phenotype/s to a playlist, make sure the phenotypes exist in the project, to list all phenotypes see `get_phenotypes`.

```{r}
name <- paste0("Diabetes", sample(1:99,1)) # Name of new playlist
descr <- "A set of all phenotypes related to diabetes in UKBB"
new_playlist <- create_playlist(name, conn, description = descr)
```

A new playlist has now been added to the project. The results come back as a `playlist` object.

```{r, echo=FALSE}
print(new_playlist)
```

## Get playlist

We can also get an existing playlist by providing the playlist's id along with the connection obect.
```{r, eval=FALSE}
pl <- get_playlist(id = 15, conn)
```

As before the results come back as a `playlist` object.

If needed, playlists can be refreshed for fetching most recent info on the playlist by calling `playlist_refresh` as follows:

```{r}
new_playlist <- playlist_refresh(new_playlist, conn)
```

## Add phenotype to playlist

Once a playlist has been fetched or created we can add phenotypes to it. As for adding phenotypes when creating a new playlist we need to make sure that the phenotype does exist in the catalog and has not been assigned to the playlist already.

```{r}
# Create a test phenotype 
pheno_name <- paste0("rtest_pheno", sample(1:99,1)) # Name of new phenotype
result_type <- "CATEGORY" # Type of phenotype (either "QT", "SET" or "CATEGORY")
tmp_phenotype <- create_phenotype(pheno_name, result_type, conn)
new_playlist  <- playlist_add_phenotype(pheno_name, new_playlist, conn)
print(new_playlist$phenotypes[1:4])
```


## Delete playlist

A playlist can easily be deleted by passing a playlist object to the `playlist_delete` function.

```{r}
playlist_delete(new_playlist, conn)
#Clean up phenotype
phenotype_delete(tmp_phenotype, conn)
```
