---
title: "Phenotype categories"
author: "Andri M. Stefansson <andri@genuitysci.com>"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Phenotype Services}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
requireNamespace("knitr")
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

In this example, we're going to explore the capabilities of the phenotype-services using the `gorr` package. 

## Load packages

First load the `gorr` package, the `tidyverse` package is recommended in general, but not required for this example

```{r packages}
library(gorr)
library(magrittr) # pipe
library(tibble)
```

## Connecting to the phenotype service

First we'll need to establish a connection to our phenotype-catalog API. To do that we'll need to call `phenotype_connect` and provide it with the relevant parameters pointing to the phenotype-catalog-service, i.e. `api_key` and `project`:


```{r}
conn <- phenotype_connect(api_key = Sys.getenv("GOR_API_KEY"),
                          project = Sys.getenv("GOR_API_PROJECT"))
conn
```

alternative way is to use the `gor_connect` and pass the relevant `api-endpoint`

```{r, eval=FALSE}
## Dont run
conn <- gor_connect(api_key = Sys.getenv("GOR_API_KEY"),
                    project = Sys.getenv("GOR_API_PROJECT"),
                    api_endpoint = "api/phenotype-catalog")
```

If everything goes as planned, we'll have a `conn` object to pass into subsequent functions.


## List categories

All available categories can be listed by passing the `conn` object to the `get_categories` function

```{r}
categories <- get_categories(conn)
sample(categories, 5)
```

The results come back as a `vector` of categories

## Create category

If we want to add a new category to the catalog we can do so using `create_category` along with the name of the category to be added as follows.

NB: Before creating a new category, make sure you have the privileges (admin rights to category management) to do so.

```{r}
name <- paste0("rtestCat", sample(1:99, 1)) # Name of new category
new_category <- create_category(name,  conn)
```

A new category has now been added and assigned to the variable `new_category`:

```{r}
print(new_category)
```