---
title: "Exploring phenotype matrix services"
author: "Andri M. Stefansson <andri@genuitysci.com>"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Phenotaype Matrix Services}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
requireNamespace("knitr")
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

In this example, we're going to explore the capabilities of the phenotype matrix services using the `gorr` package. 

## Load packages

First load the `gorr` package, the `tidyverse` package is recommended in general, but not required for this example

```{r packages}
library(gorr)
library(magrittr) # pipe
library(tibble) 
```

## Connecting to the phenotype service

First we'll need to establish a connection to our phenotype-catalog API. To do that we'll need to call `phenotype_connect` and provide it with the relevant parameters pointing to the phenotype-catalog-service, i.e. `api_key` and `project`:


```{r}
conn <- phenotype_connect(api_key = Sys.getenv("GOR_API_KEY"),
                          project = Sys.getenv("GOR_API_PROJECT"))
conn
```

alternative way is to use the `gor_connect` and pass the relevant `api-endpoint`

```{r, eval=FALSE}
# Dont run
conn <- gor_connect(api_key = Sys.getenv("GOR_API_KEY"),
                    project = Sys.getenv("GOR_API_PROJECT"),
                    api_endpoint = "api/phenotype-catalog")
```

If everything goes as planned, we'll have a `conn` object to pass into subsequent functions.

## List project phenotypes

Let's start by listing available phenotypes in project

```{r}
recipe_tag = "Phecode"
phenos<- get_phenotypes(conn, tags = recipe_tag,  limit = 3000)
phenos %>% head()
```

The results come back as a `vector` of phenotypes

```{r}
phenos_subset_names <- sample(phenos,5)
phenos_subset_names 
```

## Initialize phenotype matrix

We initialize an empty phenotype matrix object by running `get_phenotype_matrix` optionally passing the `base`variable.

```{r}
pheno_matrix <- get_phenotype_matrix()
class(pheno_matrix)
pheno_matrix
```

## Adding phenotype/s to the phenotype matrix

To add phenotype/s to the phenotype matrix we use either `phemat_add_phenotype` for adding a single phenotype to the matrix or `phemat_add_phenotypes` to add multiples.
Note that these methods do not have the same input arguments
* phemat_add_phenotype(name, phenotype_matrix, missing_value, label)
* phemat_add_phenotypes(names, phenotype_matrix, missing_value)

```{r}
pheno_matrix <- phemat_add_phenotypes(phenos_subset_names, pheno_matrix,  missing_value = '-99') 
```

```{r, echo= FALSE}
print(paste("pheno_matrix class:", class(pheno_matrix)))
print(paste("pheno_matrix content:", paste(names(pheno_matrix), collapse =  ", ")))
print(paste("pheno_matrix phenotypes:", paste(names(pheno_matrix$phenotypes), collapse =  ", ")))
```

## Remove phenotype from matrix

```{r}
rm_pheno <- pheno_matrix$phenotypes[[1]]$name
paste("Removing phenotype:", rm_pheno)
pheno_matrix <- phemat_remove_phenotype(name=rm_pheno, pheno_matrix)
pheno_matrix

```



## Get data

Lastly, let's fetch a phenotype from the project. We'll use the first listed 

```{r}
my_10_phenos <-  get_data(pheno_matrix, conn)
```

```{r}
print(my_10_phenos)
```
